\bf Comparing the Power of Plot Designs to Reveal Correlation

### 2015-12-16 paper/paper.Rnw
Updated

<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
library(lme4)
library(ggplot2)
library(mnormt)
library(tidyr)
# Convert reshape2 to tidyr, melt to gather - done
#library(reshape2)
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)
library(stringr)
#library(pander)
#library(splitstackshape) - changed cSplit to unnest
#library(readr)
#library(lubridate)
# Convert these to read_csv at some point
correl <- read.csv("correlations.csv", stringsAsFactors = FALSE)
full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) 
Turk <- read.csv("turk.csv", stringsAsFactors = FALSE)

@@

\begin{figure}[htbp]
  \centering
<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE>>=
# correl <- read.csv("correlations.csv") # You have already read this
count.setup <- select(full, pic_id, response_0, actual_location, 
              test_param, smoothed, n) #full[,c(5, 6, 10, 15, 16, 17)]
# Need to use the tidyr commands instead of cSplit
# Here's where we will need to make some changes to incorporate the multiple picks
#pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
#pin$response_0 <- separate(pin, response_0, "response_0", sep=",", extra="drop")
count.setup$choice <- separate(count.setup, response_0, c("choice", "l1", "l2"), ",", extra="drop")$choice
count.setup$choice[str_detect(count.setup$actual_location, count.setup$response_0)] <- 
  count.setup$actual_location[str_detect(count.setup$actual_location, count.setup$response_0)]
                         #as.integer(pin$response_0)
counts <- data.frame(table(count.setup$pic_id, count.setup$choice))
counts$Var1 <- as.numeric(as.character(counts$Var1))
counts$Var2 <- as.numeric(as.character(counts$Var2))
colnames(counts) <- c("lineup", "plot", "count")
pin <- merge(counts, Turk, by.x="lineup", by.y="pic_id")
pin <- select(pin, lineup, plot.x, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
pin$r <- factor(pin$r)
colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
pin$base <- 0
pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"))
pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
pin <- mutate(pin, percent = count/attempt)
pin.a <- ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
       geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
pin.b <- ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
pin.c <- ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
pin.d <- ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
grid.arrange(pin.a, pin.b, pin.c, pin.d, ncol=1)
@

@@

# Need to keep only evaluations done by subject who completed at least 10
x <- data.frame(table(full$subj_id))
keep <- x$Var1[x$Freq>9]
full.sub <- filter(full, subj_id %in% keep)
full.sub$r <- round(full.sub$correlation, 1)
full.sub$rsgn <- ifelse(full.sub$r >= 0, 1, 0)
full.sub$rabs <- abs(full.sub$r)
#full.sub$rabs <- factor(full.sub$rabs)
full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location)#[,c(1,7, 15,16,17,20:26)]
full.sub$attempts <- sapply(strsplit(as.character(full.sub$response_0), ","), length)
#full.sub$weights = 1/(full.sub$attempts) #weights are actually number of attempts in the glm
#full.sub <- full.sub %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
#full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses
#full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
#full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
full.sub$alpha <- log(.05/0.95)
full.sub$correct.prop <- full.sub$correct/full.sub$attempts #correct.prop - proportion of correct selections, weights - number of attempts

m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
m8s <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial", weights=attempts)
#pander(summary(m8l), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for overlaid lines"))
#pander(summary(m8s)$coefficients, justify=c('left', 'right', 'right', 'right', 'right'), caption="Model fit for scatter plots")
#library(xtable)
x <- summary(m8l)$coefficients
y <- summary(m8s)$coefficients
colnames(y)[4] <- "p-value"
rownames(y) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
kable(x, format="latex", digits=2)
kable(y, format="latex", digits=2)
@
\end{table}

@@

\begin{figure}[htbp]
 \centering
<<powermodel, dependson='powerfit', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE>>=
r <- seq(-0.9,0.9,0.02)
n <- c(12, 24, 48, 96)
full.pred8l <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
                                      smoothed=c(0,1), 
                                      subj_id=subset(full.sub, plot=="L")$subj_id)) #ranef(m4)[[1]][,1]))
full.pred8l$rsgn <- ifelse(full.pred8l$r > 0, 1, 0)
full.pred8l$rabs <- abs(full.pred8l$r)
full.pred8l$fitre <- predict(m8l, full.pred8l, type="response", re.form=~(1|subj_id),
                             allow.new.levels=TRUE)
full.pred8l$fitfx <- predict(m8l, full.pred8l, type="response", re.form=NA,
                             allow.new.levels=TRUE)
full.pred8l$plot <- ifelse(full.pred8l$smoothed == 0, "Overlaid Lines", "Overlaid Lines (Smoothed)")
full.pred8l <- filter(full.pred8l, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)

full.pred8s <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
                                      smoothed=c(0,1), 
                                      subj_id=subset(full.sub, plot=="S")$subj_id)) #ranef(m4)[[1]][,1]))
full.pred8s$rsgn <- ifelse(full.pred8s$r > 0, 1, 0)
full.pred8s$rabs <- abs(full.pred8s$r)
full.pred8s$fitre <- predict(m8s, full.pred8s, type="response", re.form=~(1|subj_id),
                             allow.new.levels=TRUE)
full.pred8s$fitfx <- predict(m8s, full.pred8s, type="response", re.form=NA,
                             allow.new.levels=TRUE)
full.pred8s$plot <- ifelse(full.pred8s$smoothed == 0, "Scatterplot", "Scatterplot (Smoothed)")
full.pred8s <- filter(full.pred8s, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
full.pred <- rbind(full.pred8s, full.pred8l)

@@

\begin{figure}
  \centering
<<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE>>=
resids <- resid(m8s)
residl <- resid(m8l)
resids[842:843] <- NA
resid <- data.frame(line = residl, scatter = resids)
p1b <- ggplot(data=resid[1:841,]) + geom_histogram(aes(y=..density.., x=scatter)) + labs(title="Scatter plot", x="residual", y=element_blank())
p2b <- ggplot(data=resid) + geom_histogram(aes(y=..density.., x=line))+ labs(title="Overlaid Lines", x="residual", y=element_blank())
grid.arrange(p1b, p2b, ncol=2, top = "Residual density plots")
@
  \label{residuals}
  \caption{Histograms of residuals for both the scatter plot (left) and the overlaid lines (right).}
\end{figure}

@@

<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
                  + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
kable(summary(timelmer)$coefficients, format="latex")
@

>>>>

<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
library(lme4)
library(ggplot2)
library(mnormt)
library(tidyr)
library(knitr)
# Convert reshape2 to tidyr, melt to gather - done
#library(reshape2)
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)
library(stringr)
#library(pander)
#library(splitstackshape) - changed cSplit to unnest
#library(readr)
#library(lubridate)
# Convert these to read_csv at some point
correl <- read.csv("correlations.csv", stringsAsFactors = FALSE)
full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) 
Turk <- read.csv("turk.csv", stringsAsFactors = FALSE)

@@

\begin{figure}[htbp]
  \centering
<<pins, dependson='samplesize', fig.width=6, fig.height=10, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
# correl <- read.csv("correlations.csv") # You have already read this
count.setup <- select(full, pic_id, response_0, actual_location, 
              test_param, smoothed, n)
count.setup$response_0 <- as.character(count.setup$response_0) #full[,c(5, 6, 10, 15, 16, 17)]
# Need to use the tidyr commands instead of cSplit
# Here's where we will need to make some changes to incorporate the multiple picks
#pin <- pin %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
#pin$response_0 <- separate(pin, response_0, "response_0", sep=",", extra="drop")
count.setup$choice <- separate(count.setup, response_0, c("choice", "l1", "l2"), ",", extra="drop")$choice
count.setup$choice[str_detect(count.setup$actual_location, count.setup$response_0)] <- 
  count.setup$actual_location[str_detect(count.setup$actual_location, count.setup$response_0)]
                         #as.integer(pin$response_0)
counts <- data.frame(table(count.setup$pic_id, count.setup$choice))
counts$Var1 <- as.numeric(as.character(counts$Var1))
counts$Var2 <- as.numeric(as.character(counts$Var2))
colnames(counts) <- c("lineup", "plot", "count")
pin <- merge(counts, Turk[-19], by.x="lineup", by.y="pic_id")
pin <- select(pin, lineup, plot, count, smoothed, true_pos, n, test_param, r)#pin[,c(1, 2, 3, 12, 13, 15, 11, 20)]
pin$r <- factor(pin$r)
colnames(pin) <- c("lineup", "plot", "count", "smoothed", "true_pos", "n", "design", "r")
pin <- mutate(pin, design=ifelse(design=="scatter", "Scatterplot", "Overlaid Lines"))#, smoothed=ifelse(smoothed=="no", 0, 1))
pin$base <- 0
pin$Correlation <- diag(as.matrix(correl[pin$lineup, (pin$plot+2)]))
pin <- mutate(pin, true = ifelse(plot==true_pos, "True plot", "Null plot"), percent = count/attempt)
#pin <- merge(pin, Turk[,c(2, 16)], by.x = "lineup", by.y = "pic_id")
pin.a <- ggplot(data=filter(pin, smoothed==0 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
       geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed() + facet_grid(n~r) + labs(title="Unsmoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
pin.b <- ggplot(data=subset(pin, smoothed==1 & design == "Scatterplot")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Scatterplot", y="Percent Chosen") + theme(legend.position="none")
pin.c <- ggplot(data=subset(pin, smoothed==0 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Unsmoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
pin.d <- ggplot(data=subset(pin, smoothed==1 & design == "Overlaid Lines")) + geom_errorbar(aes(x=Correlation, ymin=base, ymax=percent, colour=true)) +
  geom_point(aes(x=Correlation, y=percent, colour=true)) + coord_fixed()  + facet_grid(n~r) + labs(title="Smoothed Overlaid Lines", y="Percent Chosen") + theme(legend.position="none")
grid.arrange(pin.a, pin.b, pin.c, pin.d, ncol=1)
@

@@

# Need to keep only evaluations done by subject who completed at least 10
x <- data.frame(table(full$subj_id))
keep <- x$Var1[x$Freq>9]
full.sub <- filter(full, subj_id %in% keep)
full.sub$r <- round(full.sub$correlation, 1)
full.sub$rsgn <- ifelse(full.sub$r >= 0, 1, 0)
full.sub$rabs <- abs(full.sub$r)
#full.sub$rabs <- factor(full.sub$rabs)
full.sub$plot <- ifelse(full.sub$test_param == 0, "S", "L")
full.sub <- select(full.sub, subj_id, correct, r, rsgn, rabs, smoothed, n, plot, log_time, response_0, actual_location)#[,c(1,7, 15,16,17,20:26)]
full.sub$attempts <- sapply(strsplit(as.character(full.sub$response_0), ","), length)
full.sub$weights = 1/(full.sub$attempts) #weights are actually number of attempts in the glm
full.sub <- full.sub %>% transform(response_0= strsplit(as.character(response_0), ",")) %>% unnest(response_0)
full.sub$correct <- ifelse(full.sub$response_0 == full.sub$actual_location, 1, 0) #fixing correct for multiple guesses
#full.sub$rpos <- ifelse(full.sub$r > 0, full.sub$r, 0)
#full.sub$rneg <- ifelse(full.sub$r < 0, full.sub$r, 0)
full.sub$alpha <- log(.05/0.95)
#full.sub$correct.prop <- full.sub$correct/full.sub$attempts #correct.prop - proportion of correct selections, weights - number of attempts
#weirw.csv(full.sub, "stata.csv") #exporting data to other programs
m8l <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "L"), family="binomial", weights=attempts)
m8s <- glmer(correct.prop~offset(alpha) + rabs*rsgn + smoothed + n + (1|subj_id), data=filter(full.sub, plot == "S"), family="binomial", weights=attempts)
#pander(summary(m8l), justify=c('left', 'right', 'right', 'right', 'right', caption="Model fit for overlaid lines"))
#pander(summary(m8s)$coefficients, justify=c('left', 'right', 'right', 'right', 'right'), caption="Model fit for scatter plots")
#library(xtable)
line.fit <- summary(m8l)$coefficients
scatter.fit <- summary(m8s)$coefficients
colnames(scatter.fit)[4] <- "p-value"
rownames(scatter.fit) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
#colnames(line.fit)[4] <- "p-value"
#rownames(line.fit) <- c("Intercept", "abs(r)", "sign(r)", "smooth", "sample size", "Interaction(r+, r-)")
kable(scatter.fit, format="latex", digits=2, caption="Scatter plot")
kable(line.fit, format="latex", digits=2, caption="Overlaid Lines")
@
\end{table}

@@

\begin{figure}[htbp]
 \centering
<<powermodel, dependson='powerfit', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE>>=
r <- seq(-0.9,0.9,0.02)
n <- c(12, 24, 48, 96)
predl <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
                                      smoothed=c(0,1), 
                                      subj_id=subset(full.sub, plot=="L")$subj_id)) #ranef(m4)[[1]][,1]))
predl$rsgn <- ifelse(predl$r > 0, 1, 0)
predl$rabs <- abs(predl$r)
predl$fitre <- predict(m8l, predl, type="response", re.form=~(1|subj_id),
                             allow.new.levels=TRUE)
predl$fitfx <- predict(m8l, predl, type="response", re.form=NA,
                             allow.new.levels=TRUE)
predl$plot <- ifelse(predl$smoothed == 0, "Overlaid Lines", "Overlaid Lines (Smoothed)")
predl <- filter(predl, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)

preds <- data.frame(expand.grid(r=r, n=n, alpha=log(.05/0.95),
                                      smoothed=c(0,1), 
                                      subj_id=subset(full.sub, plot=="S")$subj_id)) #ranef(m4)[[1]][,1]))
preds$rsgn <- ifelse(preds$r > 0, 1, 0)
preds$rabs <- abs(preds$r)
preds$fitre <- predict(m8s, preds, type="response", re.form=~(1|subj_id),
                             allow.new.levels=TRUE)
preds$fitfx <- predict(m8s, preds, type="response", re.form=NA,
                             allow.new.levels=TRUE)
preds$plot <- ifelse(preds$smoothed == 0, "Scatterplot", "Scatterplot (Smoothed)")
preds <- filter(preds, smoothed == 0 | smoothed == 1 & n==48 | smoothed ==1 & n==96)
full.pred <- rbind(preds, predl)

@@

\begin{figure}
  \centering
<<residuals, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE>>=
resids <- resid(m8s)
residl <- resid(m8l)
resids[842:843] <- NA
resid <- data.frame(line = residl, scatter = resids)
p1b <- ggplot(data=resid[1:841,]) + geom_histogram(aes(y=..density.., x=scatter)) + labs(title="Scatter plot", x="residual", y=element_blank())
p2b <- ggplot(data=resid) + geom_histogram(aes(y=..density.., x=line))+ labs(title="Overlaid Lines", x="residual", y=element_blank())
grid.arrange(p1b, p2b, ncol=2, top = "Residual density plots")
@
  \label{residuals}
  \caption{Histograms of residuals for both the scatter plot (left) and the overlaid lines (right).}
\end{figure}

@@

\begin{table}
\label{tab2}
\centering
\caption{Mixed Linear Model fit for time taken to make a selection}
<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
                  + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
time.fit <- summary(timelmer)$coefficients
pv <- function(x){
  p <- 2*(1-pt(x, 1835))
}
p.value <- sapply(time.fit[,3], pv)
time.fit <- cbind(time.fit, p.value)
time.fit <- time.fit[c(1, 2, 4, 3, 5, 8, 7, 6, 9, 11, 10, 12, 14, 13),] #separate into main effects then interaction effects in same order
kable(time.fit[1:7,], format="latex", caption="Main Effects")
kable(time.fit[8:14,], format="latex", caption="Interaction Effects")
@
\end{table}
